---
title: "Soil_Index_Draft"
author: "Rose Abramoff"
date: "9/25/2017"
output: html_document
---
#Looking for gridded products to classify variables into land use type, soil type, and by climate (MAT, MAP)
##Land use type
```
MODIS IGBP land cover map (Liu et al. 2015 NCC re-sampled to 0.25◦ by dominant land use to calculate ABC per biome) 
Friedl, M. A. et al. MODIS Collection 5 global land cover: Algorithm refinements and characterization of new datasets. Remote Sens. Environ. 114, 168–182 (2010).
maybe here? need to look at Friedl paper some more: https://modis.gsfc.nasa.gov/data/
```
##Soil type
```
Soilgrids (used in Crowther et al. 2016 Nature meta-analysis as in Koven et al. 2015 Phil. Trans. R. Soc. A following the procedure in Crowther et al. 2015 Nature which seems to mean simply taking the co-located or nearest 1km resolution grid)
http://www.isric.org/explore/soilgrids
ftp://ftp.soilgrids.org/data/recent/ <- I'm having trouble connecting to this with ftp soilgrids.org
```
##Climate
```
Worldclim ("Bioclim" in Crowther et al. 2016 Nature following the procedure in Crowther et al. 2015 Nature, which seems to mean simply taking the co-located or nearest 1km resolution grid)
http://www.worldclim.org/bioclim
http://worldclim.org/version2 #can get down to 1km spatial resolution
download some 30 minute (1km resolution) data
```
#load Colin's SOC data - used with permission
```{r}
sf <- read.csv("/Users/rzabramoff/Dropbox (Climate)/External Collaborations/ISCN/Allison/SOCData_Averilletal2014Nature.csv")
```

#load Allison's N min data
```{r}
nf <- read.csv("/Users/rzabramoff/Dropbox (Climate)/External Collaborations/ISCN/Allison/Nmin_data_UpdateForRose_20170922.csv")
pf <- read.csv("/Users/rzabramoff/Dropbox (Climate)/External Collaborations/ISCN/Allison/HedleyP_database_updateforRose_20170927.csv")
```

#load WoSIS data
```{r}
df <- read.table("/Users/rzabramoff/Dropbox (Climate)/External Collaborations/ISCN/WoSIS_2016_July/wosis_201607_profiles.txt", sep="\t", header=T)
att <- read.table("/Users/rzabramoff/Dropbox (Climate)/External Collaborations/ISCN/WoSIS_2016_July/wosis_201607_attributes.txt", sep="\t", header=T)
lay <- read.table("/Users/rzabramoff/Dropbox (Climate)/External Collaborations/ISCN/WoSIS_2016_July/wosis_201607_layers.txt", sep="\t", header=T, fill=T)
```

#Make histogram plots
```{r}
#WoSIS histograms
summary(lay$orgc_value_avg)
hist(lay$orgc_value_avg, xlim = c(0,100), breaks=100, xlab=expression("SOC (gC/kg)"))
hist(lay$bdws_value_avg) #bulk density of whole soil kg/dm3
summary(lay$orgc_value_avg*lay$bdws_value_avg)

p1 <- hist(lay$orgc_value_avg*lay$bdws_value_avg,  xlim = c(0,100), breaks=1000, xlab=expression("Soil Organic Carbon (kgC"~m^3~")"), main="Soil Organic Carbon", col="dark gray") #same as kg/m3

summary(as.numeric(as.character(nf$Nmin_g_m2_yr_O_mean)))
summary(as.numeric(as.character(nf$Nmin_g_m2_yr_A_mean)))
summary(as.numeric(as.character(nf$Nmin_g_m2_yr_OA_mean))) #g/m2/horizon

p2 <- hist(as.numeric(as.character(nf$Nmin_g_m2_yr_OA_mean)), xlim = c(-10,100), breaks=200, xlab=expression("N mineralization (gN"~m^2 ~yr^-1~")"), main="Topsoil Nitrogen Mineralization", col="dark gray")

summary(lay$wrvo_value_avg)

#hist(lay$wrvo_value_avg, xlim=c(0,100), breaks=100, xlab=expression("Volumetric Water Retention ("~cm^3~"100"~cm^-3~")"))

wrvo.clean <- lay[!is.na(lay$wrvo_value_avg),]$wrvo_value_avg #remove NAs
wrvo.clean <- wrvo.clean[wrvo.clean>1]
summary(wrvo.clean)

p3 <- hist(wrvo.clean, xlim=c(0,100), breaks=50, xlab=expression("Volumetric Water Retention ("~cm^3~"100"~cm^-3~")"), main="Volumetric Water Retention", col="dark gray")
```

#Make normalized histograms
```{r}
range01 <- function(x){(x-min(x))/(max(x)-min(x))}

summary(lay$orgc_value_avg*lay$bdws_value_avg)
x = lay$orgc_value_avg*lay$bdws_value_avg  
x = x[!is.na(x)]
soc.norm = range01(x)

x = as.numeric(as.character(nf$Nmin_g_m2_yr_OA_mean))
x = x[!is.na(x)]
nmin.norm = range01(x)

x = lay$wrvo_value_avg
x = x[!is.na(x)]
wrvo.norm = range01(x)

p1 <- hist(soc.norm,  xlim = c(0,1), breaks=1000, xlab=expression("Soil Organic Carbon (kgC"~m^3~")"), main="Soil Organic Carbon", col="dark gray") #same as kg/m3

p2 <- hist(nmin.norm, xlim = c(-10,100), breaks=200, xlab=expression("N mineralization (gN"~m^2 ~yr^-1~")"), main="Topsoil Nitrogen Mineralization", col="dark gray")

p3 <- hist(wrvo.norm, xlim=c(0,100), breaks=50, xlab=expression("Volumetric Water Retention ("~cm^3~"100"~cm^-3~")"), main="Volumetric Water Retention", col="dark gray")
```
#Removing tails and suspected errors
```{r}
range01 <- function(x){(x-min(x))/(max(x)-min(x))}

x = lay$orgc_value_avg*lay$bdws_value_avg  
x = x[!is.na(x)]
x = x[x < quantile(x, 0.95) ] #remove 95% tail
soc.norm = range01(x)

x = as.numeric(as.character(nf$Nmin_g_m2_yr_OA_mean))
x = x[!is.na(x)]
x = x[x > quantile(x, 0.05) ] #remove 5% tail
x = x[x < quantile(x, 0.95) ] #remove 95% tail
nmin.norm = range01(x)

x = lay$wrvo_value_avg
x = x[!is.na(x)]
x = x[x > quantile(x, 0.83) ] #remove values lower than 0.46
wrvo.norm = range01(x)

pdf(file="/Users/rzabramoff/Dropbox (Climate)/External Collaborations/ISCN/figures/normalized_histograms.pdf")
p1 <- hist(soc.norm,  xlim = c(0,1), breaks=100, xlab=expression("Soil Organic Carbon (kgC"~m^3~")"), main="Soil Organic Carbon", col="dark gray") #same as kg/m3

p2 <- hist(nmin.norm, xlim = c(0,1), breaks=30, xlab=expression("N mineralization (gN"~m^2 ~yr^-1~")"), main="Topsoil Nitrogen Mineralization", col="dark gray")

p3 <- hist(wrvo.norm, xlim=c(0,1), breaks=100, xlab=expression("Volumetric Water Retention ("~cm^3~"100"~cm^-3~")"), main="Volumetric Water Retention", col="dark gray")
dev.off()
```
#Download RACA NRCS USDA Dataset
```{r}
raca <- read.csv(file="/Users/rzabramoff/Dropbox (Climate)/External Collaborations/ISCN/RaCA_Download/RaCA_Download/Data/RaCA_SOC_pedons.csv")

#SOC stock is in Mg/ha
head(raca)
head(raca$SOCstock100) #Mg/ha
raca$SOCkgm3 <- raca$SOCstock100/10 #convert to kg/m3
summary(raca$SOCkgm3)

hist(raca[raca$LU=="F",]$SOCkgm3)

```


#Index
```{r}
##Subsect whole data set
SOCnums <- sf[sf$LandUseType==c("forest","grass","undisturbed"),]
SOCnums <- SOCnums[!is.na(SOCnums$carbon.1m),] #remove NAs
nf$Nmin_g_m2_yr_OA_mean <- as.numeric(as.character(nf$Nmin_g_m2_yr_OA_mean))
#Nnums <- nf[nf$LandUseType==c("forest","grass","undisturbed"),]
#Nnums <- Nnums[!is.na(Nnums$Nmin_g_m2_yr_OA_mean),] #remove NAs
Nnums <- nf[!is.na(nf$Nmin_g_m2_yr_OA_mean),] #remove NAs
Nnums <- Nnums[Nnums$Nmin_g_m2_yr_OA_mean < 80 & Nnums$Nmin_g_m2_yr_OA_mean > -200,] #remove 2 outliers
Claynums <- sf[sf$LandUseType==c("forest","grass","undisturbed"),]
Claynums <- Claynums[!is.na(Claynums$clay.1m),] #remove NAs

range01 <- function(x){(x-min(x))/(max(x)-min(x))}
rangeSOC <- function(x){(x-min(SOCnums$carbon.1m))/(max(SOCnums$carbon.1m)-min(SOCnums$carbon.1m))}
rangeNmin <- function(x){(x-min(Nnums$Nmin_g_m2_yr_OA_mean))/(max(Nnums$Nmin_g_m2_yr_OA_mean)-min(Nnums$Nmin_g_m2_yr_OA_mean))}
rangeClay <- function(x){(x-min(Claynums$clay.1m))/(max(Claynums$clay.1m)-min(Claynums$clay.1m))}

W.Index <- function(s,n,c,ws,wn,wc){
SOCindex <- rangeSOC(EnterYourSOCValue)
Nindex <- rangeNmin(EnterYourNminValue)
Clayindex <- rangeClay(EnterYourClayValue)
windex <- (SOCindex*ws + Nindex*wn + Clayindex*wc)/3
return(windex)
}

socs <- NULL
nmins <- NULL
clays <- NULL

socs[[1]] = tapply(SOCnums$carbon.1m, SOCnums$LandUseType, FUN=min)
nmins[[1]] = tapply(Nnums$Nmin_g_m2_yr_OA_mean, Nnums$LandUseType, FUN=min, na.rm=T)
clays[[1]] = tapply(Claynums$clay.1m, Claynums$LandUseType, FUN=min, na.rm=T)
socs[[2]] = tapply(SOCnums$carbon.1m, SOCnums$LandUseType, FUN=mean)
nmins[[2]] = tapply(Nnums$Nmin_g_m2_yr_OA_mean, Nnums$LandUseType, FUN=mean, na.rm=T)
clays[[2]] = tapply(Claynums$clay.1m, Claynums$LandUseType, FUN=mean, na.rm=T)
socs[[3]] = tapply(SOCnums$carbon.1m, SOCnums$LandUseType, FUN=max)
nmins[[3]] = tapply(Nnums$Nmin_g_m2_yr_OA_mean, Nnums$LandUseType, FUN=max, na.rm=T)
clays[[3]] = tapply(Claynums$clay.1m, Claynums$LandUseType, FUN=max, na.rm=T)

Equalweight=NULL
SOCweighted=NULL
Clayweighted=NULL
Nminweighted=NULL

for (i in 1:3){
EnterYourSOCValue <- socs[[i]][c("forest","grass","undisturbed")] #kgC/m2 to 1m
EnterYourNminValue <- nmins[[i]][c("forest","grass","undisturbed")] #in g/m2/yr for OA horizons
EnterYourClayValue <- clays[[i]][c("forest","grass","undisturbed")] #in %
SOCWeight <- 1
NminWeight <- 1
ClayWeight <- 1

Equalweight[[i]] <- W.Index(EnterYourSOCValue,EnterYourNminValue,EnterYourClayValue,SOCWeight,NminWeight,ClayWeight)

SOCWeight <- 1.5
NminWeight <- 0.5
ClayWeight <- 0.5

SOCweighted[[i]] <- W.Index(EnterYourSOCValue,EnterYourNminValue,EnterYourClayValue,SOCWeight,NminWeight,ClayWeight)

SOCWeight <- 0.5
NminWeight <- 0.5
ClayWeight <- 1.5

Clayweighted[[i]] <- W.Index(EnterYourSOCValue,EnterYourNminValue,EnterYourClayValue,SOCWeight,NminWeight,ClayWeight)

SOCWeight <- 0.5
NminWeight <- 1.5
ClayWeight <- 0.5

Nminweighted[[i]] <- W.Index(EnterYourSOCValue,EnterYourNminValue,EnterYourClayValue,SOCWeight,NminWeight,ClayWeight)
}

#pdf(file="/Users/rzabramoff/Dropbox (Climate)/External Collaborations/ISCN/figures/mean_index_by_landuse_barplots.pdf")
tab = rbind(Equalweight[[2]],SOCweighted[[2]])
tabmin = rbind(Equalweight[[1]],SOCweighted[[1]])
tabmax = rbind(Equalweight[[3]],SOCweighted[[3]])
barplot(tab, beside = T, legend=T, legend.text = c("Equal Weight","SOC Weighted 3x"), ylim=c(0,0.3))
segments(tab, c(tabmin[1,],tabmin[2,]), tab, c(tabmax[1,],tabmax[2,]), lwd = 1.5, col=1) #!
arrows(tab, tabmin, tab, tabmax, lwd = 1.5, angle = 90, code = 3, length = 0.05) #!

tab2 = rbind(Equalweight[[2]],SOCweighted[[2]], Nminweighted[[2]], Clayweighted[[2]])
barplot(tab2, beside = T, col= c(1,"brown","dark green","gray"), legend=T, legend.text = c("Equal Weight","SOC Weighted 3x","Nmin Weighted 3x","Clay Weighted 3x"), ylim=c(0,1))
#dev.off()

#Example
means <- c(23, 28, 19)
names <- c("squirrel", "rabbit", "chipmunk")
standardErrors <- c(1.2, 1.7, 0.9)
plotTop <- max(means+standardErrors*2)
barCenters <- barplot(means, names.arg=names, col="gray", las=1, ylim=c(0,plotTop))
segments(barCenters, means-standardErrors*2, barCenters, means+standardErrors*2, lwd=2)
```


#Preliminary Index from 0-1
```{r}
EnterYourSOCValue <- 100 #in gC/dm3 - same as kg/m3
EnterYourNminValue <- 30 #in g/m2/yr for OA horizons
EnterYourClayValue <- 60 #in %
SOCWeight <- 1
NminWeight <- 1
ClayWeight <- 1

W.Index(EnterYourSOCValue,EnterYourNminValue,EnterYourClayValue,SOCWeight,NminWeight,ClayWeight)
```


#Other scripts
#Make plots
```{r}
library(ggplot2)

#SOC plot
p.soc <- ggplot(SOCnums, aes(carbon.1m, fill=LandUseType)) +
  stat_density(position="identity", alpha=0.7) +
  xlab("Soil C (kgC/m2) - 1m") +
  theme(text = element_text(size=20))

#N min plot
p.nmin <- ggplot(Nnums, aes(Nmin_g_m2_yr_OA_mean, fill=LandUseType)) +
  stat_density(position="identity", alpha=0.7) +
  #xlim(-10,60) +
  xlab("N mineralization (g/m2/y) - OA horizon") +
  theme(text = element_text(size=20))

#Clay plot
p.clay <- ggplot(Claynums, aes(clay.1m, fill=LandUseType)) +
  stat_density(position="identity", alpha=0.7) +
  xlab("Clay content (%) - 1m") +
  theme(text = element_text(size=20))

pdf(file="/Users/rzabramoff/Dropbox (Climate)/External Collaborations/ISCN/density_plots.pdf")
p.soc
p.nmin
p.clay
dev.off()
```

#Make normalized plots
```{r}
pdf(file="/Users/rzabramoff/Dropbox (Climate)/External Collaborations/ISCN/density_plots_normalized.pdf")

#SOC plot
SOCnums$norm.carbon.1m <- rangeSOC(SOCnums$carbon.1m)
ggplot(SOCnums, aes(norm.carbon.1m, fill=LandUseType)) +
  stat_density(position="identity", alpha=0.7) +
  xlab("Soil C (normalized) - 1m") +
  theme(text = element_text(size=20)) +
  xlim(0,1)

#N min plot
Nnums$norm.NminNum <- rangeNmin(Nnums$Nmin_g_m2_yr_OA_mean)
ggplot(Nnums, aes(norm.NminNum, fill=LandUseType)) +
  stat_density(position="identity", alpha=0.7) +
  #xlim(-10,60) +
  xlab("N mineralization (normalized) - OA horizon") +
  theme(text = element_text(size=20)) +
  xlim(0,1)

#Clay plot
Claynums$norm.clay.1m <- rangeClay(Claynums$clay.1m)
ggplot(Claynums, aes(norm.clay.1m, fill=LandUseType)) +
  stat_density(position="identity", alpha=0.7) +
  xlab("Clay content (normalized) - 1m") +
  theme(text = element_text(size=20))+
  xlim(0,1)

dev.off()
```

#Map of sites - find sites that overlap in the two datasets
```{r}
library(ggplot2)
library(ggmap)
library(maps)
library(mapdata)
w2hr <- map_data("world2Hires")
wrld <- map_data("world")
usa <- map_data("usa")
sf$lat <- as.numeric(sf$latitude)
sf$long <- as.numeric(sf$longitude)
nf$lat <- as.numeric(as.character(nf$Latitude))
nf$long <- as.numeric(as.character(nf$Longitude))
```

#Global map
```{r}
gg1 <- ggplot() + 
  geom_polygon(data = wrld, aes(x=long, y = lat, group = group), fill = NA, color = "black") + 
  coord_fixed(1.3)
#pdf(file="/Users/rzabramoff/Dropbox (Climate)/External Collaborations/ISCN/sites_map_world.pdf")
gg1 + 
  geom_point(data = sf, aes(x = long, y = lat), color = "red", size = 1) +
  geom_point(data = nf, aes(x = long, y = lat), color = "blue", size = 1)
#dev.off()
```

#USA
```{r}
gg2 <- ggplot() + 
  geom_polygon(data = usa, aes(x=long, y = lat, group = group), fill = NA, color = "black") + 
  coord_fixed(1.3) +
  ylim(25,50) +
  xlim(-130,-65)
#pdf(file="/Users/rzabramoff/Dropbox (Climate)/External Collaborations/ISCN/sites_map_usa.pdf")
gg2 + 
  geom_point(data = sf, aes(x = long, y = lat), color = "red", size = 3) +
  geom_point(data = nf, aes(x = long, y = lat), color = "blue", size = 3)
#dev.off()
```


